### 2018-12-03 Revision/Response/response.tex

\noindent A large gap in the eigenvalue spectrum typically indicates that a low-dimensional (1--2) active subspace 
would sufficiently capture the variability in the model output due to uncertainty in the inputs. However,
in the revised manuscript, we have shown that the variability in the ignition delay is reasonably
approximated with a 1-dimensional active subspace even though the gap in the first two eigenvalues
is found to be O(10). Hence, a smaller gap in the eigenvalue spectrum might still lead to a reasonable
accuracy with a low-dimensional active subspace depending upon the relationship between the
system response and the inputs. However, it is also possible that we might need to a consider a higher 
dimensional active subspace to be able to approximate the variability in the response with reasonable
accuracy in situations where the gap is small.

\item \referee{11.33: Alg. 1: should have $\beta$ as input, too?}

\noindent The reviewer is correct. Algorithm 1 in the revised manuscript has been updated to include $\beta$
as an input.

\item \referee{11.42: Which norm is used? Any intuition why squares of W are used, and not just W?}

\noindent We have used the L$^2$ norm to estimate the relative change in the squared values of corresponding
eigenvector components between successive iterations. The algorithm in the revised manuscript has been updated 
to specify the norm. 
We consider square of individual components due to the fact the both W and -W contain legitimate eigenvectors 
of C. Alternatively, we could have also used the modulus value of the components to assess convergence.

\item \referee{11.45: what is max over?}

\noindent The max is over all eigenvectors in the active subspace denoted by $j$. This has been specified in
the updated algorithm in the revised manuscript.

\item \referee{12.29: 'higher-order' should be replaced with higher-index or smaller-magnitude, to avoid confusion.}

\noindent The reviewer's suggestion has been incorporated in the revised manuscript.

### 2018-12-04 Revision/Response/response.tex

\baselineskip=22pt

\vspace*{1in}

%We thank the reviewers for their assessment of our manuscript. Please find
%below point-by-point replies (in black) to your comments and questions
%(reprinted in blue). Where possible, key modifications have been highlighted
%in blue in the revised manuscript. We sincerely hope that with these
%modifications, the paper is found suitable for publication in the
%{\it Journal of Scientific Computing}.

We are extremely thankful to the reviewers for recognizing potential in this effort and 
for their constructive feedback on the manuscript. A point-by-point
response to the comments has been provided below. Corresponding modifications have been
highlighted (in blue) in the revised manuscript. Major changes in the revised manuscript
are summarized as follows:

\begin{itemize}

\item The complexity of the problem has been increased substantially in two ways: First, 
the non-linearity of the system response has been signficantly increased by increasing
the perturbations associated with the uncertain rate-controlling parameters: pre-exponent
and the activation energy. Second, in addition to the rate parameters, the impact of
perturbing the initial conditions: pressure, temperature, and stoichiometry pertaining to the
H$_2$/O$_2$ reaction mechanism on the ignition delay is investigated. Such analysis using the
proposed framework would provide useful insights to combustion scientists interested in
kinetics applications. Moreover, it illustrates the applicability and robustness of the 
framework to complex kinetics applications.

\item The two approaches, perturbation and regression, for estimating the model gradient
with respect to uncertain parameters and inputs are compared for convergence behavior and
predictive accuracy using the same amount of
computational effort in the revised manuscript, as suggested by one of the reviewers.

\end{itemize}

We sincerely hope that based on the modifications in the revised manuscript and our 
response to reviewers' comments, the paper is found suitable for publication in
{\it Combustion and Flame}.

@@

\noindent The max is over all eigenvectors in the active subspace denoted by $j$. This has been specified in
the updated algorithm in the revised manuscript.

\item \referee{12.29: 'higher-order' should be replaced with higher-index or smaller-magnitude, to avoid confusion.}

\noindent The reviewer's suggestion has been incorporated in the revised manuscript.

\item \referee{Figs 2a, 3 and 4 demonstrate the regression-based gradient computation is working
similarly to the perturbation based one. However, the comparison is really qualitative and does
 not provide much insight. Clearly, regression will approximate the gradient less accurately then
a local perturbation, but the latter has a price associated with it as new simulations are needed
(a factor of d+1). It would have been great to see how accurate each method is for comparable
expense, or vice-verse, how expensive each approach is for a comparable accuracy. I know there
is a statement about this at the end of Section 4, but it is neither sufficient nor quantitative.
Same goes for statements in Section 5.2 - it was not clear whether regression- and
perturbation- based approaches had comparable expenses (in terms of model evaluations) or not.}

\noindent Based on the reviewer's feedback, in the revised manuscript, the analysis has been
updated to
compare the perturbation and regression approaches for their convergence and predictive 
accuracy using the same amount of computational effort. As a result, the comparative 
analysis is much more conclusive with regards to the performance of the two approaches.

\item \referee{For Figure 8b, the reasoning why E$_15$ was not captured by the regression approach,
is not satisfactory. I think the authors could have gone a bit deeper in order to explain this
glaring miss of the regression-based approach. There is some discussion in Section 6, but it is
equally unsatisfactory.}

\noindent Our latest analysis in the revised manuscript has revealed that the two approaches
in fact yield consistent results
for sensitivity analysis even for the 36-dimensional case with increased non-linearity
when using the same amount of
computational effort or the same number of model evaluations. Earlier, the two approaches were
compared for the same number of samples, N. However, the number of model runs in the case of
perturbation were N(d+1); d is the number of uncertain parameters and inputs to the model. Whereas,
the number of model runs in the case of regression were only N. Therefore, it can be said that
the accuracy of the regression-based approximation of the gradient can be increased to the 
extent of perturbation-based approach by increasing the sample size.

\item \referee{14.4-8: the near-linearity simply makes the problem extremely benign, nearly any
reasonable method would have captured such behavior. It would have been great to consider more
complicated case (more than simply going from 19 to 33 parameters), perhaps looking at logA 
perturbation instead of A would help? Also consider more than 1$\%$ 
perturbations to the input. Otherwise, the problem is too simple and does not do justice to the method.}

\noindent We have made a sincere effort to incorporate these valuable suggestions by the 
reviewer in the revised manuscript. 
Specifically, we perturb $\log(A)$ instead of $A$ which is observed to significantly increase
the non-linearity of the system response. In addition the rate parameters, we analyse the impact
of perturbing the initial conditions. Moreover, as suggested by the reviewer, we have
increased the mangnitude of the perturbations from 1$\%$ to 2$\%$ to illustrate the applicability
and robustness of the proposed approach to more complex scenarios.

\item \referee{14.8: here and in a few other places, I have an issue with statements like
active subspace is 1-dimensional'. To me, active subspace is as many dimensional as one likes,
so the statement should be changed to something like '1-dimensional active subspace captures
most of the variance/uncertainty/dynamics etc..', or '1-dimensional active subspace provides
very accurate approximation...'.}

\item \referee{14.57: even if nominal is zero, one could perturb them, right? 
Or it is actually indicating the fact that there should be no exponential factor whatsoever.}

\noindent Based on our conversations with the developers of the TChem software at Sandia
National Laboratories, we decided not to perturb the activation energies with a 0
nominal value.

\item \referee{18.35: the remark in the parentheses did not make sense to me.}

\noindent We have provided examples for perturbation-based techniques such as finite difference and
adjoint-based methods for estimating the
gradient of the model output with respect to uncertain parameters and inputs.

### 2018-12-06 Revision/Paper/method.tex
..

\bigskip
\begin{breakablealgorithm}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
  \caption{An iterative strategy for discovering the active subspace}
  \begin{algorithmic}[1]
\Require $\theta_l$, $\theta_u$, $\tau$. 
\Ensure $\hat{\mat{\Lambda}}$, $\hat{\mat{W}}$, $\bm{\nu}_r(f)$ %$\eta$. %
    \Procedure{Active Subspace Computation}{}
    \State Set $k$ = 0
	\State Draw $n_k$ random samples, $\{\bm{\xi}_i\}_{i=1}^{n_k}$ 
         according to $\pi_{\bm{\xi}}$. 
    \State Set $N_\text{total}$ = $n_k$ 
	\iffalse \State Project to the physical space:
        $\{\bm{\theta}_k\}_{k=1}^{n_r}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\xi}_k\}_{k=1}^{n_r}$ \fi
	\State For each $i=1, \ldots, N_\text{total}$, compute $f(\bm{\xi}_i)$ and the gradient $\bm{g}^i = \nabla_{\bm{\xi}}f(\bm{\xi}_i)$
    \iffalse  
	\Statex\hspace{5mm} Using Finite Difference: 
	\Statex\hspace{5mm} i. Assign a small increment, $d\xi$.
	\Statex\hspace{5mm} ii. Augment the set of samples with neighboring points.
	\be \{\bm{\Xi}_k\}_{k=1}^{n_1(N_p+1)}:~\{\bm{\xi}_k\}_{k=1}^{n_1} \cup
        \{\xi_{k,j}+d\xi\}_{j=1}^{N_p} \nonumber
	\ee
	\Statex\hspace{5mm} iii. Project to the physical space:
        $\{\bm{\theta}_k\}_{k=1}^{n_1(N_p+1)}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\Xi}_k\}_{k=1}^{n_1(N_p+1)}$
	\Statex\hspace{5mm} iv. Using the augmented set, $\{\bm{\theta}_k\}_{k=1}^{n_1(N_p+1)}$
         compute $\bm{g}^i$.
         \fi 
	\State Compute $\hat{\mat{C}}$ and its eigenvalue decomposition 
		$\hat{\mat{C}}$= $\frac{1}{N_\text{total}}\sum\limits_{i=1}^{N_\text{total}}[\bm{g}^i][\bm{g}^i]^\top$ = 
		$\hat{\mat{W}}^{(k)}\hat{\mat{\Lambda}}^{(k)} \hat{\mat{W}}^{(k)\top}$
	%\State Eigenvalue decomposition, $\mat{C}$ = $W^{(0)}\Lambda^{(0)} W^{(0)\top}$%
	\State Partition: $\hat{\mat{\Lambda}}^{(k)}=
        \begin{bmatrix} \hat{\mat{\Lambda}}_1^{(k)} & \\ & \hat{\mat{\Lambda}}_2^{(k)} \end{bmatrix}$, 
        $\hat{\mat{W}}^{(k)}=\begin{bmatrix} \hat{\mat{W}}_1^{(k)} & \hat{\mat{W}}_2^{(k)} \end{bmatrix}$, 
        $\hat{\mat{\Lambda}}_1^{(k)}\in \mathbb{R}^{N_p\times r}$
	\Loop
		\State Set $k$ = $k$ + 1
		\State Draw $n_k =  \lceil\beta n_{k-1}\rceil$  new random samples 
                $\{\bm{\xi}_i\}_{i=1}^{n_k}$  $\beta\in[0,1]$

	%	\State Project $\bm{\xi}_k$~$\rightarrow$~$\bm{\theta}_k$.%
		\State Set $N_\text{total}$ = $N_\text{total}$ + $n_k$ 
		\State Compute $\bm{g}^i = \nabla_{\bm{\xi}_i}f(\bm{\xi}_i)$, 
             	$i=n_{k-1}+1, \ldots, n_{k-1}+n_k$.  
		\State Compute $\hat{\mat{C}}$ = 
        	$\frac{1}{N_\text{total}}\sum\limits_{k=1}^{N_\text{total}}[\bm{g}^i][\bm{g}^i]^\top$
		\State Eigenvalue decomposition, $\hat{\mat{C}}$ = $\hat{\mat{W}}^{(k)}\hat{\mat{\Lambda}}^{(k)}
		 \hat{\mat{W}}^{(k)\top}$
		\State Partition the eigenspace of $\hat{\mat{C}}$ as shown in Step 7
		\State Compute $\delta \hat{\mat{W}}_{1,j}^{(k)}$ = 
                       \scalebox{1.25}{$\frac{\|(\hat{\mat{W}}_{1,j}^{k})^2 - 
                       (\hat{\mat{W}}_{1,j}^{k-1})^2\|}{\|(\hat{\mat{W}}_{1,j}^{k-1})^2\|}$}, 
                       $j = 1,\ldots,r$.
		\If {$\max\left(\delta \hat{\mat{W}}_{1,j}^{(k)}\right)<\tau$}
			\State break
		\EndIf
	\EndLoop
	\State Compute $\nu_{i,r}(f) = \sum\limits_{j=1}^{r} \lambda_j w_{i,j}^2$,
	$i=1,\ldots,N_p$.
	\State Normalize $\nu_{i,r}(f)$ as $\tilde{\nu}_{i,r}(f)$ = \scalebox{1.25}{$\frac{\nu_{i,r}(f)}{\sum_i\nu_{i,r}(f)}$}.

@@

To assess its feasibility and suitability, we implement
Algorithm~\ref{alg:grad} to compute the active subspace for 
the 19-dimensional H$_2$/O$_2$ reaction kinetics
problem by perturbing \rebut{$\log(A_i)$ by 3$\%$ about its nominal
value as discussed earlier}. For the purpose of verification,
$\hat{\mat{C}}$ was initially constructed using a large set of samples
($N$~=~1000) in the input domain. The gradient was estimated using
finite difference, and hence, a total of 20,000 model runs were performed. 
In Figure~\ref{fig:eig_comp}, we illustrate the comparison of
the resulting normalized eigenvalue spectrum \rebut{by plotting $\log(\lambda_i/\lambda_0)$} ($i = 1,\ldots,19$) corresponding to $N$~=~1000 and the same quantity corresponding to
a much smaller set of samples, $n$~=~$\{20,40,80,120\}$.
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.45\textwidth]{./Figures/eig_comp_p3}
%   \includegraphics[width=0.48\textwidth]{./Figures/err_eigv_1}
\caption{A comparison of the normalized eigenvalue spectrum, $\log(\lambda_i/\lambda_0)$
using $n$~=~$\{20,40,80,120\}$ samples with that
obtained using a much larger sample size, $N$~=~1000. 
%Right: Relative $L_2$ 
%norm of the difference between
%the square of individual components ($w_i$'s) of the dominant eigenvector evaluated using 
%$N$=1000 and $n$~=~$\{20,40,80,120\}$ samples.
} 
\label{fig:eig_comp}
\end{center}
\end{figure}
%
We observe that the dominant eigenvalues, $\lambda_1, \ldots, \lambda_4$, 
are approximated 
reasonably well with just 20 samples. As expected, the accuracy of higher-order eigenvalues is observed
to improve with the sample size. Since 
$\lambda_1$ is roughly \rebut{an order of magnitude} larger than $\lambda_2$, we expect 
a 1-dimensional active subspace \rebut{to reasonably approximate the uncertainty
in the ignition delay}. 
To further confirm this, we evaluate a relative L$^2$ norm of the difference
 ($\varepsilon_{\text{L}^{2}}^{N-n}$) between the 
squared value of corresponding components of the dominant eigenvector, computed using $N$~=~1000 ($\vec{w}_{1,N}$)
and $n$~=~$\{20,40,80,120\}$ ($\vec{w}_{1,n}$) as follows:
%
\be
\varepsilon_{\text{L}^{2}}^{N-n} = 
\frac{\|\vec{w}_{1,N}^2-\vec{w}_{1,n}^2\|_2}{\|\vec{w}_{1,N}^2\|_2}
\label{eq:accu}
\ee
%
The quantity, $\varepsilon_{\text{L}^{2}}^{N-n}$, was found to be 
$\mathcal{O}(10^{-2})$ in all cases.
Thus, even a small sample size, $n$ = 20, seems to approximate the dominant eigenspace with
reasonable accuracy in this case. 
The iterative strategy therefore offers a significant potential for computational gains.

@@

\rebut{The quantity, $\max(\delta \hat{\mat{W}}_{1,j}^{(r)})$ is plotted in
Figure~\ref{fig:comp} (center) to illustrate convergence characteristics based on Algorithm~\ref{alg:grad} 
for the two approaches. As observed for a set tolerance $\tau$~=~0.05, 8 iterations are needed 
to converge in both cases. 
At each iteration, 5 new random samples were added for the perturbation approach. As discussed earlier, gradient 
computation (using finite difference) at 5 samples requires 100 model runs. Hence, in order to compare the
two approaches for the same computational effort, we generate 100 random samples in the 19D input domain for 
the regression approach. Therefore, 8 iterations require 800 model runs leading to a total of 900 model runs 
including the initial step in both cases.  The accuracy of the two approaches is compared for the same 
computational effort by computing $\varepsilon_{\text{L}^{2}}^{N-n}$ using~\eqref{eq:accu} in 
Figure~\ref{fig:comp} (right). Once again, $N$
corresponds to the case of active subspace computation using the perturbation approach using 1000 samples
in the 19D domain; $n$ ranges from 5 to 45 for the perturbation approach, and 100 to 900 for the regression 
approach corresponding to iterations 1 to 8. The accuracy of the two approaches is found to be comparable for
the same amount of computational effort. However, the perturbation approach is observed to be slightly more
accurate.}
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.8\textwidth]{./Figures/eig_convD19}
\caption{\rebut{Left: An illustrative comparison of individual squared components of the converged dominant
eigenvector obtained using perturbation and regression strategies using $M$~=~900 model runs in each case. 
Additionally, the dominant eigenvector components obtained using $M$~=~20000 model runs (corresponding to 
$N$~=~1000 samples)
in the perturbation strategy (test case), used to assess the accuracy of the two strategies are also plotted.
Center: The quantity,  $\max(\delta \hat{\mat{W}}_{1,j}^{(r)})$
is plotted for successive iterations to illustrate convergence behavior for the two strategies. Right: The
quantity, $\varepsilon_{\text{L}^{2}}^{N-n}$ is plotted as a function of iterations to compare the accuracy
of the two strategies. Note that the above plots compare the convergence and accuracy of the two
strategies for the same amount of computational effort, quantified by the number of model runs, $M$.}}
\label{fig:comp}
\end{center}
\end{figure}
\rebut{In Figure~\ref{fig:comp} (left), we compare the
squared values of individual components of the dominant eigenvector obtained using the two approaches
at the end of 8 iterations in each case. Additionally, the dominant eigenvector components obtained
for a test case involving $N$~=~1000 (and $M$ = 20,000 model runs)
samples using the perturbation approach are also plotted. 
The set of eigenvector components for the three cases are found to be in excellent agreement with
each other, indicating that both
approaches have sufficiently converged and are reasonably accurate for this setup.}

>>>>

\bigskip
\begin{breakablealgorithm}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
  \caption{An iterative strategy for discovering the active subspace}
  \begin{algorithmic}[1]
\Require $\theta_l$, $\theta_u$, \rebut{$\beta$}, $\tau$. 
\Ensure $\hat{\mat{\Lambda}}$, $\hat{\mat{W}}$, $\bm{\nu}_r(f)$ %$\eta$. %
    \Procedure{Active Subspace Computation}{}
    \State Set $k$ = 0
	\State Draw $n_k$ random samples, $\{\bm{\xi}_i\}_{i=1}^{n_k}$ 
         according to $\pi_{\bm{\xi}}$. 
    \State Set $N_\text{total}$ = $n_k$ 
	\iffalse \State Project to the physical space:
        $\{\bm{\theta}_k\}_{k=1}^{n_r}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\xi}_k\}_{k=1}^{n_r}$ \fi
	\State For each $i=1, \ldots, N_\text{total}$, compute $f(\bm{\xi}_i)$ and the gradient $\bm{g}^i = \nabla_{\bm{\xi}}f(\bm{\xi}_i)$
    \iffalse  
	\Statex\hspace{5mm} Using Finite Difference: 
	\Statex\hspace{5mm} i. Assign a small increment, $d\xi$.
	\Statex\hspace{5mm} ii. Augment the set of samples with neighboring points.
	\be \{\bm{\Xi}_k\}_{k=1}^{n_1(N_p+1)}:~\{\bm{\xi}_k\}_{k=1}^{n_1} \cup
        \{\xi_{k,j}+d\xi\}_{j=1}^{N_p} \nonumber
	\ee
	\Statex\hspace{5mm} iii. Project to the physical space:
        $\{\bm{\theta}_k\}_{k=1}^{n_1(N_p+1)}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\Xi}_k\}_{k=1}^{n_1(N_p+1)}$
	\Statex\hspace{5mm} iv. Using the augmented set, $\{\bm{\theta}_k\}_{k=1}^{n_1(N_p+1)}$
         compute $\bm{g}^i$.
         \fi 
	\State Compute $\hat{\mat{C}}$ and its eigenvalue decomposition 
		$\hat{\mat{C}}$= $\frac{1}{N_\text{total}}\sum\limits_{i=1}^{N_\text{total}}[\bm{g}^i][\bm{g}^i]^\top$ = 
		$\hat{\mat{W}}^{(k)}\hat{\mat{\Lambda}}^{(k)} \hat{\mat{W}}^{(k)\top}$
	%\State Eigenvalue decomposition, $\mat{C}$ = $W^{(0)}\Lambda^{(0)} W^{(0)\top}$%
	\State Partition: $\hat{\mat{\Lambda}}^{(k)}=
        \begin{bmatrix} \hat{\mat{\Lambda}}_1^{(k)} & \\ & \hat{\mat{\Lambda}}_2^{(k)} \end{bmatrix}$, 
        $\hat{\mat{W}}^{(k)}=\begin{bmatrix} \hat{\mat{W}}_1^{(k)} & \hat{\mat{W}}_2^{(k)} \end{bmatrix}$, 
        $\hat{\mat{\Lambda}}_1^{(k)}\in \mathbb{R}^{N_p\times r}$
	\Loop
		\State Set $k$ = $k$ + 1
		\State Draw $n_k =  \lceil\beta n_{k-1}\rceil$  new random samples 
                $\{\bm{\xi}_i\}_{i=1}^{n_k}$  $\beta\in[0,1]$

	%	\State Project $\bm{\xi}_k$~$\rightarrow$~$\bm{\theta}_k$.%
		\State Set $N_\text{total}$ = $N_\text{total}$ + $n_k$ 
		\State Compute $\bm{g}^i = \nabla_{\bm{\xi}_i}f(\bm{\xi}_i)$, 
             	$i=n_{k-1}+1, \ldots, n_{k-1}+n_k$.  
		\State Compute $\hat{\mat{C}}$ = 
        	$\frac{1}{N_\text{total}}\sum\limits_{k=1}^{N_\text{total}}[\bm{g}^i][\bm{g}^i]^\top$
		\State Eigenvalue decomposition, $\hat{\mat{C}}$ = $\hat{\mat{W}}^{(k)}\hat{\mat{\Lambda}}^{(k)}
		 \hat{\mat{W}}^{(k)\top}$
		\State Partition the eigenspace of $\hat{\mat{C}}$ as shown in Step 7
		\State Compute $\delta \hat{\mat{W}}_{1,j}^{(k)}$ = 
                       \scalebox{1.25}{$\frac{\|(\hat{\mat{W}}_{1,j}^{k})^2 - 
                       (\hat{\mat{W}}_{1,j}^{k-1})^2\|_\rebut{2}}{\|(\hat{\mat{W}}_{1,j}^{k-1})^2\|_\rebut{2}}$}, 
                       $j = 1,\ldots,r$.
		\If {$\max\limits_{\rebut{j}}\left(\delta \hat{\mat{W}}_{1,j}^{(k)}\right)<\tau$}
			\State break
		\EndIf
	\EndLoop
	\State Compute $\nu_{i,r}(f) = \sum\limits_{j=1}^{r} \lambda_j w_{i,j}^2$,
	$i=1,\ldots,N_p$.
	\State Normalize $\nu_{i,r}(f)$ as $\tilde{\nu}_{i,r}(f)$ = \scalebox{1.25}{$\frac{\nu_{i,r}(f)}{\sum_i\nu_{i,r}(f)}$}.

@@

To assess its feasibility and suitability, we implement
Algorithm~\ref{alg:grad} to compute the active subspace for 
the 19-dimensional H$_2$/O$_2$ reaction kinetics
problem by perturbing \rebut{$\log(A_i)$ by 3$\%$ about its nominal
value as discussed earlier}. For the purpose of verification,
$\hat{\mat{C}}$ was initially constructed using a large set of samples
($N$~=~1000) in the input domain. The gradient was estimated using
finite difference, and hence, a total of 20,000 model runs were performed. 
In Figure~\ref{fig:eig_comp}, we illustrate the comparison of
the resulting normalized eigenvalue spectrum \rebut{by plotting $\log(\lambda_i/\lambda_0)$} ($i = 1,\ldots,19$) corresponding to $N$~=~1000 and the same quantity corresponding to
a much smaller set of samples, $n$~=~$\{20,40,80,120\}$.
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.45\textwidth]{./Figures/eig_comp_p3}
%   \includegraphics[width=0.48\textwidth]{./Figures/err_eigv_1}
\caption{A comparison of the normalized eigenvalue spectrum, $\log(\lambda_i/\lambda_0)$
using $n$~=~$\{20,40,80,120\}$ samples with that
obtained using a much larger sample size, $N$~=~1000. 
%Right: Relative $L_2$ 
%norm of the difference between
%the square of individual components ($w_i$'s) of the dominant eigenvector evaluated using 
%$N$=1000 and $n$~=~$\{20,40,80,120\}$ samples.
} 
\label{fig:eig_comp}
\end{center}
\end{figure}
%
We observe that the dominant eigenvalues, $\lambda_1, \ldots, \lambda_4$, 
are approximated 
reasonably well with just 20 samples. As expected, the accuracy of higher-\rebut{index}
 eigenvalues is observed
to improve with the sample size. Since 
$\lambda_1$ is roughly \rebut{an order of magnitude} larger than $\lambda_2$, we expect 
a 1-dimensional active subspace \rebut{to reasonably approximate the uncertainty
in the ignition delay}. 
To further confirm this, we evaluate a relative L$^2$ norm of the difference
 ($\varepsilon_{\text{L}^{2}}^{N-n}$) between the 
squared value of corresponding components of the dominant eigenvector, computed using $N$~=~1000 ($\vec{w}_{1,N}$)
and $n$~=~$\{20,40,80,120\}$ ($\vec{w}_{1,n}$) as follows:
%
\be
\varepsilon_{\text{L}^{2}}^{N-n} = 
\frac{\|\vec{w}_{1,N}^2-\vec{w}_{1,n}^2\|_2}{\|\vec{w}_{1,N}^2\|_2}
\label{eq:accu}
\ee
%
The quantity, $\varepsilon_{\text{L}^{2}}^{N-n}$, was found to be 
$\mathcal{O}(10^{-2})$ in all cases.
Thus, even a small sample size, $n$ = 20, seems to approximate the dominant eigenspace with
reasonable accuracy in this case. 
The iterative strategy therefore offers a significant potential for computational gains.

@@

\rebut{The quantity, $\max(\delta \hat{\mat{W}}_{1,j}^{(r)})$ defined in Algorithm~\ref{alg:grad} was
used to assess the convergence behavior of the two approaches. Using a set tolerance, $\tau$~=~0.05,
it was observed that both, perturbation and regression approaches took 8 iterations to converge. 
Note that the computational effort at each iteration was considered to be the same in both cases.
More specifically, 5 new random samples were added for the perturbation approach at each iteration.
However, as discussed earlier, as total of 100 (=5$\times$(19+1)) model runs were needed to obtain the
model prediction and its gradients at these newly generated samples. Hence, in the case of
regression, 100 new random samples were generated at each iteration since gradient computation does
not require additional model runs in this case. Thus, including the initial step, a total of 900
model runs were required to obtained a converged active subspace in both cases.

The accuracy of the two approaches was assessed by estimating $\varepsilon_{\text{L}^{2}}^{N-n}$ 
using the components of the dominant eigenvector in the converged active subspace in each case
in~\eqref{eq:accu}. The quantity, $\varepsilon_{\text{L}^{2}}^{N-n}$ was estimated to be 0.0657
and 0.1050 using perturbation and regression respectively. Hence, the regression approach was
found to be relatively more accurate. Squared values of the individual components of the 
dominant eigenvector from the
two approaches and for the case using $N$~=~1000 in the perturbation approach are plotted in
Figure~\ref{fig:comp}.}
%
%is plotted in
%Figure~\ref{fig:comp} (center) to illustrate convergence characteristics based on Algorithm~\ref{alg:grad} 
%for the two approaches. As observed for a set tolerance $\tau$~=~0.05, 8 iterations are needed 
%to converge in both cases. 
%At each iteration, 5 new random samples were added for the perturbation approach. As discussed earlier, gradient 
%computation (using finite difference) at 5 samples requires 100 model runs. Hence, in order to compare the
%two approaches for the same computational effort, we generate 100 random samples in the 19D input domain for 
%the regression approach. Therefore, 8 iterations require 800 model runs leading to a total of 900 model runs 
%including the initial step in both cases.  The accuracy of the two approaches is compared for the same 
%computational effort by computing $\varepsilon_{\text{L}^{2}}^{N-n}$ using~\eqref{eq:accu} in 
%Figure~\ref{fig:comp} (right). Once again, $N$
%corresponds to the case of active subspace computation using the perturbation approach using 1000 samples
%in the 19D domain; $n$ ranges from 5 to 45 for the perturbation approach, and 100 to 900 for the regression 
%approach corresponding to iterations 1 to 8. The accuracy of the two approaches is found to be comparable for
%the same amount of computational effort. However, the perturbation approach is observed to be slightly more
%accurate.}
%
%\begin{figure}[htbp]
% \begin{center}
%  \includegraphics[width=0.8\textwidth]{./Figures/eig_convD19}
%\caption{\rebut{Left: An illustrative comparison of individual squared components of the converged dominant
%eigenvector obtained using perturbation and regression strategies using $M$~=~900 model runs in each case. 
%Additionally, the dominant eigenvector components obtained using $M$~=~20000 model runs (corresponding to 
%$N$~=~1000 samples)
%in the perturbation strategy (test case), used to assess the accuracy of the two strategies are also plotted.
%Center: The quantity,  $\max(\delta \hat{\mat{W}}_{1,j}^{(r)})$
%is plotted for successive iterations to illustrate convergence behavior for the two strategies. Right: The
%quantity, $\varepsilon_{\text{L}^{2}}^{N-n}$ is plotted as a function of iterations to compare the accuracy
%of the two strategies. Note that the above plots compare the convergence and accuracy of the two
%strategies for the same amount of computational effort, quantified by the number of model runs, $M$.}}
%\label{fig:comp}
%\end{center}
%\end{figure}
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.45\textwidth]{./Figures/eig_conv19D}
\caption{\rebut{An illustrative comparison of individual squared components of the converged dominant
eigenvector obtained using perturbation and regression strategies using $M$~=~900 model runs in each case. 
Additionally, the dominant eigenvector components obtained using $M$~=~20000 model runs (corresponding to 
$N$~=~1000 samples)
in the perturbation strategy (test case), used to assess the accuracy of the two strategies are also plotted.}}
\label{fig:comp}
\end{center}
\end{figure}
%
\rebut{
The set of eigenvector components for the three cases are found to be in excellent agreement with
each other, indicating that both
approaches have sufficiently converged and are reasonably accurate for this setup.}
